\chapter {Materials and Methods}

The objective of this study was to measure and quantify MRI-induce translational and torque forces experienced by an  post-surgical epicardial pacing leads when exposed to the static magnetic field. Safety limits for MRI magnetic fields are well established and described in the Introduction, based on ASTM F-series guidelines. Following these methods, we constructed non-magnetic test apparatus specifically designed to measure these forces with the precision required for clinical applications.

However, clinical access for experimental testing posed a significant challenge, as MRI scanner time is typically associated with long scheduling delays and short available measurement periods. Therefore, a secondary objective of this work was to improve the quality of our experimental setup in order to maximize the efficiency of data acquisition during these limited observation windows.

Previous students developed the laboratory-scale instruments and associated software necessary to evaluate magnetic forces in a controlled environment. This was achieved by generating a smaller but reproducible magnetic field in the laboratory, enabling characterization of materials with relatively high magnetic susceptibility. While this laboratory-generated field was not sufficient to directly measure or characterize the epicardial lead itself, it allowed for evaluation of the accuracy and measurement capacity of each instrument, as well as for the observation of the physical phenomena described above. Furthermore, given the proportional nature of the magnetic effects on a material, it was expected that, when tested in the much stronger (3 T) magnetic field of a clinical MRI system, the small magnetic susceptibility of the epicardial lead would not present a significant limitation in terms of measurement accuracy.

Three distinct experimental setups were involved in this study: (1) a spatial field mapping platform to characterize $B_0$ and $\nabla B_0$, (2) a pendulum-based deflection apparatus to assess translational force through angular displacement, and (3) a torque measurement system for rotational force assessments. In this chapter we will explain our setup and rationale, as well as the proven modifications to the instruments in order to provide a reproducible map of action to test our instruments.

\input{MatsNMets/Magnetic Field Generation}

\input{MatsNMets/Translational}

\input{MatsNMets/Torque}
